---
title: "Initial Korean covid 19"
author: "Areli Nava"
date: "6/21/23"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   


  
# Your Question - Is dying independent of sex?

## Methods

We will analyzed the sex of the patients and the result if they are alive or dead. I will the variables in my collection of Cat~cat. This is a type of problem that we can use that we can incorporate the helper file and copy into our project to gather the correct answer.








## Graphical results - use ggplot commands to do this  

#### Plot barplots ggplot way
```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


In the counting chart I can see that not a lot of people people died. It is 50-70. In the percentage in the chart female is 50% is alive and male percentage is about 40%. there is more female than male, twice as many males were dying.





  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
In the percentage of the chart I can observe that more female are alive with the difference of 12. 36% of dead are female and 64% are male. In the second chart females are 98.7 and males are only 97. In the dead percentage females are 1.4% and males are 3%.


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In the group of 1831 female alive and 1411 male were alive.in retrospect females are more alive than males in percentage. The p value is 1 out of a thousand in the Chi-square, In Fisher's test 9 out of 10,000 that are about the same p value. there is and odd ratio in the chart. The odds of females living against the odds of males dying. We want to be aware of the odds that are occurring.

## Conclusions

In todays lesson I learned to to make a ggplot with the helper file, I applied results and changed the file project to kRdata. I created a numerical graph in percent and counting dead or alive. I also ran a chi-square test and Fisher's test and how both are equal in p value. In the final result we wanted to know the odds in the results of female and male casualties, we found men suffered more during the study while females had a more better outcome.

